Genetic alterations and expression patterns of genes coding for FGF, FGFR, and FGFBP proteins in urothelial cancers

### 2024-12-17 report/markdown/report.Rmd
Additional literature references
References for the Leiden algorithm and the imputation method used in the BCAN expression data set

The BCAN study information was extracted from [cBioportal](https://www.cbioportal.org/) and consisted of demographic and clinical information, staging of primary tumors, overall and disease-specific survival, RNA sequencing expression Z scores of bulk cancer samples, as well as detailed somatic mutation data of bulk cancer specimens. 
This collective consisted of metastatic cancers, treated in most patients by systemic chemotherapy. 
The biological material was collected both in urogenital tract and at distant metastatic sites.

@@

Co-expression networks of the FGF-, FGFR-, and FGFBP-coding genes were constructed in form of weighted graphs for pairwise correlations of the expression levels with Spearman's $\rho \geq$ 0.3 (function `as_iGraph()`, package [_graphExtra_](https://github.com/PiotrTymoszuk/graphExtra) with wrapper tools for _igraph_) [@Csardi2006]. 
Isolated vertices were removed with `prune_degree()` function. 
The following vertex importance statistics [@Csardi2006] were computed with `summary()` method called for the network objects: 
degree (number of directly connected neighbor nodes), 
hub score (eigenvector of the affinity matrix, i.e. an overall strength of correlations), and 
betweenness (number of the shortest paths between all pairs of vertices that pass through the vertex). 
The networks were visualized with Fruchterman - Reingold algorithm by calling `plot()` method [@Briatte2021].

>>>>

The BCAN study information was extracted from [cBioportal](https://www.cbioportal.org/) and consisted of demographic and clinical information, staging of primary tumors, overall and disease-specific survival, RNA sequencing expression Z scores of bulk cancer samples, as well as detailed somatic mutation data of bulk cancer specimens. 
This collective consisted of metastatic cancers, treated in most patients by systemic chemotherapy. 
The biological material was collected both in urogenital tract and at distant metastatic sites. 
Missing values in the Z-score expression data set of the BCAN cohort were imputed with the 9-nearest neighbor algorithm implemented by function `impute.knn()` from _impute_ package [@Hastie2023].

@@

Co-expression networks of the FGF-, FGFR-, and FGFBP-coding genes were constructed in form of weighted graphs for pairwise correlations of the expression levels with Spearman's $\rho \geq$ 0.3 (function `as_iGraph()`, package [_graphExtra_](https://github.com/PiotrTymoszuk/graphExtra) with wrapper tools for _igraph_) [@Csardi2006]. 
Isolated vertices were removed with `prune_degree()` function. 
Communities of co-regulated genes in the co-expression network in the DepMap urothelial cancer cell line collective were identified with the Leiden algorithm (loss function: modularity, resolution: $\gamma$ = 0.5, iterations: n = 100, function `cluster_leiden()`, _igraph_ package) [@Csardi2006; @Traag2019].

The following vertex importance statistics [@Csardi2006] were computed with `summary()` method called for entire network objects (bulk cancer cohorts) or gene community sub-networks (DepMap cell lines): 
degree (number of directly connected neighbor nodes), 
hub score (eigenvector of the affinity matrix, i.e. an overall strength of correlations), and 
betweenness (number of the shortest paths between all pairs of vertices that pass through the vertex). 
The networks were visualized with Fruchterman - Reingold algorithm by calling `plot()` method [@Briatte2021].
